---
title: 'CMTH 642 Data Analytics: Advanced Methods'

---
<center> <h1> Assignment 3 (10%) </h1> </center>
<center>  <h3> Mohsen Qureshi </h2> </center>
---


#### 1. Import to R the following file: http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv (The dataset is related to white Portuguese "Vinho Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality) (3 points)
```{r}
Wine_Database <- read.csv(file = "http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", head = TRUE, sep=";")
```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
str(Wine_Database)
Wine_Database[, 1:11] <- sapply(Wine_Database[, 1:11],
as.numeric)
str(Wine_Database)
```
  
#### 3. Are there any missing values in the dataset? (4 points)
```{r}
sum(is.na(Wine_Database) == TRUE)
```

#### 4. What is the correlation between the attributes other than Quality? (10 points)
```{r}
Wine_correlation <- cor(Wine_Database[-12])
Wine_correlation
install.packages("PerformanceAnalytics",repos = "http://cran.us.r-project.org")
library("PerformanceAnalytics")
chart.Correlation(Wine_Database[-12], histogram = TRUE)
```

#### 5. Graph the frequency distribution of wine quality by using Quality. (10 points)
```{r}
hist(Wine_Database$quality,
col = "Red",
main = "Wine Quality",
xlab = "Score")
```

#### 6.	Reduce the levels of rating for quality to three levels as high, medium and low. Assign the levels of 3 and 4 to level 0; 5 and 6 to level 1; and 7,8 and 9 to level 2. (10 points)
```{r}
Wine_Database$quality <- ifelse(Wine_Database$quality < 5, 'Low',
ifelse(Wine_Database$quality > 7, 'High', 'Medium'))
Wine_Database$quality <- ordered(Wine_Database$quality, c('Low',
'Medium', 'High'))
round(prop.table(table(Wine_Database$quality)) * 100, digits = 1)
```

#### 7. Normalize the data set by using the following function: (12 points)
```{r}
normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
N_Wine_Database <- as.data.frame(sapply(Wine_Database[-12], normalize))
```

#### 8. Divide the dataset to training and test sets. (12 points)
```{r}
Index <- sample(nrow(N_Wine_Database), floor(0.65 * nrow(N_Wine_Database)))
WineQuality_Train <- N_Wine_Database[Index,]
WineQuality_Test <- N_Wine_Database[-Index,]
WineQuality_Train_Labels <- Wine_Database[Index, 12]
WineQuality_Test_Labels <- Wine_Database[-Index, 12]
```

#### 9. Use the KNN algorithm to predict the quality of wine using its attributes. (12 points)
```{r}
install.packages("class", repos = "http://cran.us.r-project.org")
library(class)
install.packages("gmodels", repos = "http://cran.us.r-project.org")
library(gmodels)

Wine_Prediction <- knn(WineQuality_Train[1:11],
WineQuality_Test[1:11], WineQuality_Train_Labels, 7)
head(Wine_Prediction)

```

#### 10. Display the confusion matrix to evaluate the model performance. (12 points)
```{r}
Wine_Quality_CM <- table(Actual = WineQuality_Test_Labels,
Predicted = Wine_Prediction)
Wine_Quality_CM

CrossTable(x = WineQuality_Test_Labels, y = Wine_Prediction,prop.chisq = F)
```


This is the end of Assignment 3
